---
title: "Weather Is Not Climate"
author: "Russell Szumski"
date: 2017-12-20
categories: ["R"]
tags: ["ggplot2", "ggiraph"]
---


I live in Chicago and it's cold old, but weather is not climate. The difference between weather and climate is a measure of time. Weather is what conditions of the atmosphere are over a short period of time, and climate is how the atmosphere "behaves" over relatively long periods of time.

Let's take the data and plot the deviations from the 1951 - 1980 average going back to 1880 and see what we get.


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)
library(ggplot2)
library(ggthemes)
library(ggiraph)

options(warn=-1)

df_temp = read.csv('Annual Temp Dev/ZonAnn.Ts+dSST.csv')

pal <- c("#313695","#4575b4","#74add1","#abd9e9","#e0f3f8","#ffffbf","#fee090","#fdae61","#f46d43","#d73027","#a50026")
vals <- seq(-2, 2, length = 10)

df_temp$tooltip <- paste0(df_temp$Year,':', df_temp$Glob)

tooltip_css <- "background-color:black;color:white;padding:10px;"

plot_int <- ggplot(df_temp, aes(x = Year, y = Glob, frame = Year, tooltip = tooltip)) +
  geom_point_interactive(shape = 21, colour="black", aes(fill = Glob), size=5, stroke=1) +
  scale_x_continuous(limits=c(1880,2017)) +
  scale_y_continuous(limits=c(-0.4,1.2)) +
  theme_minimal() +
  scale_fill_gradientn(colors = pal, values = vals, rescaler = function(x, ...) x, oob = identity, guide=FALSE) +
  labs(x = "Year",
       y = "Deviations from 1951-1980 (C)",
       title = "Land-Ocean: Global Annual Means ",
       subtitle = "") +
  geom_hline(yintercept = 0) +
  theme_minimal(base_family = "RobotoCondensed-Regular") +
  theme(text=element_text(size=12, family="Roboto-Bold"))

p <- ggiraph(code = print(plot_int),hover_css = "fill-opacity:5",
        tooltip_extra_css = tooltip_css, width = .9, height = 4) 

```


```{r, echo=FALSE}
p
```

While I might be experiencing some cold weather, the average global temperature is on the rise. Sixteen of the 17 warmest years in the 136-year record all have occurred since 2001 and 2016 ranks as the warmest on record.

Tools: R

Data Source: [NASA](https://climate.nasa.gov/system/internal_resources/details/original/647_Global_Temperature_Data_File.txt)