---
title: "Chicago Homicide Clearance Rate"
author: "Russell Szumski"
date: 2018-07-20
categories: ["R"]
tags: ["QGIS", "Mapbox", "Chart.js"]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(htmltools)
```

If you're murdered in America, there's a 1 in 3 chance that the police won't identify your killer.

To use the FBI's terminology, the national "clearance rate" for homicide today is 64.1 percent. Fifty years ago, it was more than 90 percent.

From 2001 to 2017, of the 8913 homicides in Chicago, the average percent of homicides that resulted in an arrest was just under 49. The number of homicides has spiked in recent years while the arrest rate decreased. There have been 4609 homicides without an arrest between 2001 and 2017.

```{r,echo=F}
htmltools::tags$iframe(src = 'https://rszumski.github.io/chicago_homicides_bar_charts/', width = '100%', height = '400', style="border:none;")
```

Chicago has low-arrest zones with a high concentration of homicides and low arrest rates, and high-arrest zones with a high concentration of homicides and high arrest rates.

```{r,echo=F}
htmltools::tags$iframe(src = 'https://rszumski.github.io/chicago_homicides/', width = '100%', height = '800', style="border:none;")
```

Tools:Mapbox, Chart.js, R, QGIS <br> Data Source: [Chicago Data Portal](https://data.cityofchicago.org/Public-Safety/Crimes-2001-to-present/ijzp-q8t2)

