---
title: "Chicago Homicide Clearance Rate"
author: "Russell Szumski"
date: 2018-07-20
categories: ["R"]
tags: ["highcharter", "flexdashboard", "leaflet"]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(htmltools)
```

If you're murdered in America, there's a 1 in 3 chance that the police won't identify your killer.

To use the FBI's terminology, the national "clearance rate" for homicide today is 64.1 percent. Fifty years ago, it was more than 90 percent.

From 2001 to 2017, of the 8913 homicides in Chicago, the average percent of homicides that resulted in an arrest was just under 49. The number of homicides has spiked in recent years while the arrest rate decreased. There have been 4609 homicides without an arrest between 2001 and 2017.

```{r,echo=F}
htmltools::tags$iframe(src = 'https://rszumski.github.io/chicago_homicides_bar_charts/', width = 100%, height = '600', style="border:none;")
```
Chicago has low-arrest zones with a high concentration of homicides and low arrest rates, and high-arrest zones with a high concentration of homicides and high arrest rates.

```{r,echo=F}
htmltools::tags$iframe(src = 'https://rszumski.github.io/chicago_homicides/', width = 100%, height = '800', style="border:none;")
```

A [link](https://rstudio-pubs-static.s3.amazonaws.com/406148_53cd0545428c46c3a573c2538d86ed1c.html) to a full page view.



I followed Illinois Uniform Crime Reporting (IUCR) codes when filtering out homicides from other crimes. IUCR codes are four digit codes that law enforcement agencies use to classify criminal incidents when taking individual reports. In that program, homicides include FIRST DEGREE MURDER, SECOND DEGREE MURDER, INVOLUNTARY MANSLAUGHTER, and RECKLESS HOMICIDE.

Clearance rate or arrest rate is calculated by dividing the number of crimes that are "cleared" (a charge being laid, an arrest) by the total number of crimes recorded.

#### Mapping Methodology

Areas shaded in orange are places where fewer than one-third of the homicides resulted in an arrest. 

Areas shaded in blue are where two-thirds or more of the homicides resulted in an arrest
