---
title: "Chicago Homicide Clearance Rate"
author: "Russell Szumski"
date: 2018-07-20
categories: ["R"]
tags: ["highcharter", "flexdashboard", "leaflet"]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(htmltools)
```

If you're murdered in America, there's a 1 in 3 chance that the police won't identify your killer.

To use the FBI's terminology, the national "clearance rate" for homicide today is 64.1 percent. Fifty years ago, it was more than 90 percent.

Living in Chicago I often hear, from people outside of the city, how dangerous it is and how many murders go unsolved. So I thought I would take a look at the city's "clearance" rate for myself. Luckily for me, the city of Chicago has a publicly available dataset of crimes in the city from 2001 to the present. Using that I put together this dashboard.

```{r,echo=F}
htmltools::tags$iframe(src = 'https://rstudio-pubs-static.s3.amazonaws.com/406148_53cd0545428c46c3a573c2538d86ed1c.html', width = '1000', height = '800', style="border:none;")
```

A [link](https://rstudio-pubs-static.s3.amazonaws.com/406148_53cd0545428c46c3a573c2538d86ed1c.html) to a full page view.

#### The Data

The dataset comes from the [Chicago Data Potal](https://data.cityofchicago.org/Public-Safety/Crimes-2001-to-present/ijzp-q8t2).
It reflects reported incidents of crime (with the exception of murders where data exists for each victim) that occurred in the City of Chicago from 2001 to 2017. Data is extracted from the Chicago Police Department's CLEAR (Citizen Law Enforcement Analysis and Reporting) system. 

Disclaimer: These crimes may be based upon preliminary information supplied to the Police Department by the reporting parties that have not been verified. The preliminary crime classifications may be changed at a later date based upon additional investigation and there is always the possibility of mechanical or human error. Therefore, the Chicago Police Department does not guarantee (either expressed or implied) the accuracy, completeness, timeliness, or correct sequencing of the information and the information should not be used for comparison purposes over time.

#### Definitions

I followed Illinois Uniform Crime Reporting (IUCR) codes when filtering out homicides from other crimes. IUCR codes are four digit codes that law enforcement agencies use to classify criminal incidents when taking individual reports. In that program, homicides include FIRST DEGREE MURDER, SECOND DEGREE MURDER, INVOLUNTARY MANSLAUGHTER, and RECKLESS HOMICIDE.

Clearance rate or arrest rate is calculated by dividing the number of crimes that are "cleared" (a charge being laid, an arrest) by the total number of crimes recorded.

#### Mapping Methodology

Areas shaded in orange are places where fewer than one-third of the homicides resulted in an arrest. 

Areas shaded in blue are where two-thirds or more of the homicides resulted in an arrest