---
title: "Sad Songs"
author: "Russell Szumski"
date: 2017-12-20
categories: ["R"]
tags: ["spotifyr", "ggplot2", "ggiraph"]
---

Tools: R

Data Source: Spotify

[Gregory Alan Isakov](https://gregoryalanisakov.com/) has been one of my favorite singer-songwriters for a while, and I've gotten used to people suggesting that I play something 'less depressing'. Most singer-songwriter's music is indeed sad, but his music can't be that sad. Can it?

Using the spotifyr package we can pull track audio features from the 'Spotify' Web API, including valence and energy of a track.

Valence: a measure from 0.0 to 1.0 describing the musical positiveness conveyed by a track. Tracks with high valence sound more positive (e.g. happy, cheerful, euphoric), while tracks with low valence sound more negative (e.g. sad, depressed, angry).

Energy: a measure from 0.0 to 1.0 and represents a perceptual measure of intensity and activity. Typically, energetic tracks feel fast, loud, and noisy. For example, death metal has high energy, while a Bach prelude scores low on the scale.

Plotting valence vs. energy we can get an idea of the emotion of a song. 


```{r setup, include=FALSE, cache = TRUE}
library(tidyverse)
library(ggplot2)
library(ggiraph)
library(ggthemes)

spotify_df <- read.csv('Gregory Alan/spotify_df.csv')

spotify_df$track_name <- gsub("'", '', spotify_df$track_name)
spotify_df$album_name <- gsub("'", '', spotify_df$album_name)

spotify_df$tooltip <- paste0('<a style = "margin-right:55px">',
                             '<img src=', spotify_df$album_img, ' height="75" style="float:left;margin-right:5px">',
                             '<b>Album:</b> ', spotify_df$album_name,
                             '<br><b>Track:</b> ', spotify_df$track_name,
                             '<br><b>Valence:</b> ', round(spotify_df$valence, 2),
                             '<br><b>Energy:</b> ', round(spotify_df$energy, 2),
                             '</a>')

plot <- ggplot(spotify_df, aes(valence, energy, color = album_name, tooltip = tooltip)) +
  geom_point_interactive() +
  theme_hc(bgcolor = "sandsignika") +
  scale_colour_solarized("red") +
  geom_vline(xintercept = 0.5) +
  geom_hline(yintercept = 0.5) +
  scale_x_continuous(limits = c(0, 1)) +
  scale_y_continuous(limits = c(0, 1)) +
  guides(col = guide_legend(ncol = 2)) +
  annotate("text", x = 0.08, y = 0, label = "Sad/Depressing",  colour = "black") +
  annotate("text", x = 0.08, y = 1, label = "Turbulent/Angry",  colour = "black") +
  annotate("text", x = 0.93, y = 1, label = "Happy/Joyful",  colour = "black") +
  annotate("text", x = 0.93, y = 0, label = "Chill/Peaceful",  colour = "black") +
  theme(legend.position = "bottom",
        legend.title = element_blank()) +
  labs(x = "Valence",
       y = "Energy",
       title = " ",
       subtitle = "")
```



```{r , echo=FALSE}
ggiraph(code = {print(plot)},tooltip_offx = 20, tooltip_offy = -10)
```

Turns out yes. Of all five albums, and 55 songs,  Greg only has three songs that fall out of the sad/depressing area, with nothing landing in the happy/joyful area. 