---
title: "Instant Ramen"
author: "Russell Szumski"
date: 2018-01-10
categories: ["R"]
tags: ["tidytext", "network3D"]
---

A dataset of 2682 Instant Noodle Ratings was poseted on the [datasets](https://www.reddit.com/r/datasets/) subreddit. So I created bi-grams to see how often word X is followed by word Y with names of instant ramen noodles to find any flavors and key attributes/phrases used in the names that are grouped together and appear most often. Why? Why not, who hasn't enjoyed some instant ramen before.

Used [tidytext](https://github.com/juliasilge/tidytext) for the calculations and [networkD3](https://github.com/christophergandrud/networkD3) for the visualization.


```{r setup, include=FALSE, cache = TRUE}
library(dplyr)
library(tidytext)
library(tidyr)
library(igraph)
library(ggraph)

df <- read.csv('Ramen/The Ramen Rater - The Big List.csv')


ramen_bigrams <- df %>%
  unnest_tokens(bigram, Name, token = "ngrams", n = 2)


bigrams_separated <- ramen_bigrams %>%
  separate(bigram, c("word1", "word2"), sep = " ")

bigrams_filtered <- bigrams_separated %>%
  filter(!word1 %in% stop_words$word) %>%
  filter(!word2 %in% stop_words$word)

# new bigram counts:
bigram_counts <- bigrams_filtered %>% 
  count(word1, word2, sort = TRUE)

bigram_graph <- bigram_counts %>%
  filter(n > 20) %>%
  graph_from_data_frame()

ggraph(bigram_graph, layout = "fr") +
  geom_edge_link() +
  geom_node_point() +
  geom_node_text(aes(label = name), vjust = 1, hjust = 1)


wc <- cluster_walktrap(bigram_graph)
members <- membership(wc)


bigram_D3 <- igraph_to_networkD3(bigram_graph, group = members)

```

```{r , echo=FALSE}
head(bigram_counts, 15)
```

Looks like ramen names contain 'instant' and 'noodles' frequently, 'cup' is included a lot, and 'chicken' is a popular flavor, followed by 'beef'.

Let's visualize the bi-grams.


```{r , echo=FALSE}
library(networkD3)

forceNetwork(Links = bigram_D3$links, Nodes = bigram_D3$nodes, 
             Source = "source", Target = "target",
             Value = "value", NodeID = "name",
             Group = "group", opacity = 1.25, zoom = TRUE, bounded = TRUE,
             linkWidth = 1,
             colourScale = JS("d3.scaleOrdinal(d3.schemeCategory10);"))
```

We can see the common "blueprint"" of words in instant ramen names are focused around 'noodle', 'noodles', 'instant', 'flavor', and 'chicken'.

Tools: R

Data Source: [2682 Instant Ramen Noodle Ratings](https://www.theramenrater.com/resources-2/the-list/)
