---
title: "I Ran a Marathon Once"
author: "Russell Szumski"
date: 2018-05-20
categories: ["R"]
tags: ["highcharter", "flexdashboard", "dygraphs"]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(htmltools)
library(dplyr)
df <- read.csv('Running/Running_Stats.csv')

```

In 2016 I trained for and ran the Chicago Marathon. During the training I was wearing a Garmin Forerunner10 watch. I recorded all the stats the watch had, then did absolutely nothing with that data. It's two years later, so I guess I should take a look at it. 

When I trained I didn't follow any training program. My idea was to run a lot, I think I did that, but at this point I have no Idea. So I thought it might be fun to make a simple dashboard and have a look at what my "training" looked like. Here it is


```{r,echo=F}
htmltools::tags$iframe(src = 'https://rstudio-pubs-static.s3.amazonaws.com/393898_6dbfd3c7ebe948a6b2cfa26bec00510c.html', width = '1000', height = '800', style="border:none;")
```

A [link](https://rstudio-pubs-static.s3.amazonaws.com/393898_6dbfd3c7ebe948a6b2cfa26bec00510c.html) to a full page view.

A few things that stuck out:

-Average runs were only 5.8 miles  
-My longest run was about 18 miles and I did 53 days before the actual marathon  
-I ran for a total of 3 Days, 14 Hours, 15 Minutes and 36 Seconds  
-I ran the distance (562.33 miles) from Texarkana,TX to Atlanta,GA  
-August was my busiest month  
-Wednesday was my favorite day to run  
-My pace didn't really change  
-63830 food calories seems like a lot but is apparently equivalent to about 2.2 gallons of gas  

