---
title: " I Want to Believe"
author: "Russell Szumski"
date: 2018-01-17
categories: ["R"]
tags: ["dygraphs", "threejs"]
---


I've been watching the 11th season of The X-Files and it got me thinking about UFOs. In my search for data on UFO sightings I came across a dataset on Kaggle that has 80,000+ documented close encounters from the past 70 years. I also came across a R package called [dygraphs](https://rstudio.github.io/dygraphs/), for charting time-series data, and the [three.js](https://github.com/bwlewis/rthreejs) package, that has globejs: a somewhat silly widget that plots data and images on a 3-d globe. So let's put this all together. 


```{r setup, include=FALSE, cache = FALSE}
library(dplyr)
library(lubridate)
library(tidyr)
library(threejs)
library(dygraphs)

df <- read.csv('UFO Sightings/ufo_sighting_data.csv')

df$latitude <- as.numeric(as.character(df$latitude))

df_ <- df %>%
  mutate(latitude = round(latitude, 2), longitude = round(longitude, 2)) %>%
  .[,10:11] %>%
  group_by(latitude, longitude) %>%
  summarize(count = n()) %>%
  na.omit() 


df_$q <- as.numeric(
  cut(df_$count,
      breaks=quantile(df_$count, probs=c(0,0.90,0.95,0.99,1)),
      include.lowest=TRUE))

col = c("#0055ff","#00aaff","#00ffaa","#aaff00")[df_$q]

globe <- globejs(lat=df_$latitude, long=df_$longitude,
        val = df_$q^3,
        color=col,
        pointsize=0.5,
        atmosphere=TRUE)

df_country <- read.csv('UFO Sightings/ufo per country.csv')

ts_us <- ts(df_country$us, start = 1940, end = 2014)
ts_gb <- ts(df_country$gb, start = 1940, end = 2014)
ts_de <- ts(df_country$de, start = 1940, end = 2014)
ts_ca <- ts(df_country$ca, start = 1940, end = 2014)
ts_au <- ts(df_country$au, start = 1940, end = 2014)
ts_unknown <- ts(df_country$unknown, start = 1940, end = 2014)

UFO <- cbind(ts_au, ts_ca, ts_de, ts_gb, ts_unknown, ts_us)

dy_ts <- dygraph(UFO, main = 'The Truth is Out There') %>%
  dyOptions(colors = RColorBrewer::brewer.pal(6, 'Dark2'),
            fillGraph = T, fillAlpha = 0.4) %>%
  dyHighlight(highlightCircleSize = 4,
              hideOnMouseOut = F,
              highlightSeriesOpts = list(strokeWidth = 3)) %>%
  dyAxis("x", label = 'Year', drawGrid = F) %>%
  dyAxis("y", label = 'UFO Sightings') %>%
  dySeries("ts_au", label = 'Australia') %>%
  dySeries("ts_ca", label = 'Canada') %>%
  dySeries("ts_de", label = 'Germany') %>%
  dySeries("ts_gb", label = 'United Kingdom') %>%
  dySeries("ts_unknown", label = 'Unknown') %>%
  dySeries("ts_us", label = 'United States') %>%
  dyLegend(width = 650) %>%
  dyRangeSelector(dateWindow = c('1950-1-1', '2014-1-1')) %>%
  dyEvent("1993-09-10", 'The X-Files Starts', 'bottom') %>%
  dyAnnotation('1947-1-1', text = 'R', tooltip = 'Trust no one')

```

First a little data preparation, then let's count the occurrences for each country

```{r , eval=FALSE}
df$Date_time <- as.POSIXct(df$Date_time,format="%m/%d/%Y %H:%M")
df$Date_time <- as.Date(df$Date_time)
df$Date_time <- year(df$Date_time)

df_country <- df %>%
  group_by(Date_time,country) %>% 
  summarize(count=n())

df_country <- df_country %>% 
  spread(country, count) %>%
  replace(is.na(.), 0)

colnames(df_country)[2] <- "unknown"

df_country <- df_country %>%
  filter(date >= "1940")
```

dygraphs plots xts time series objects so lets transform the data in time series and use cbind the combined time series for each country into one time series object.

```{r , eval=FALSE}
ts_us <- ts(df_country$us, start = 1940, end = 2014)
ts_gb <- ts(df_country$gb, start = 1940, end = 2014)
ts_de <- ts(df_country$de, start = 1940, end = 2014)
ts_ca <- ts(df_country$ca, start = 1940, end = 2014)
ts_au <- ts(df_country$au, start = 1940, end = 2014)
ts_unknown <- ts(df_country$unknown, start = 1940, end = 2014)

UFO <- cbind(ts_au, ts_ca, ts_de, ts_gb, ts_unknown, ts_us)
```

Finally the plot

```{r , echo=FALSE}
dy_ts
```

We can see a large increase of sightings at the beginning of the 90's, when The X-Files started, and that most of the sightings occur in the US.

Coincidence?

Next let's make that somewhat silly 3-d globe widget.

Again a little data preparation, then let's count the occurrences for groups of latitude and longitude.

```{r , eval=FALSE}
df$latitude <- as.numeric(as.character(df$latitude))

df_ <- df %>%
  mutate(latitude = round(latitude, 2), longitude = round(longitude, 2)) %>%
  .[,10:11] %>%
  group_by(latitude, longitude) %>%
  summarize(count = n()) %>%
  na.omit() 

# Cut the values up into levels corresponding to the
# 99th, 95th, 90th, percentiles and then all the rest

df_$q <- as.numeric(
  cut(df_$count,
      breaks=quantile(df_$count, probs=c(0,0.90,0.95,0.99,1)),
      include.lowest=TRUE))

# Colors for each level

col = c("#0055ff","#00aaff","#00ffaa","#aaff00")[df_$q]
```

The widget.

```{r , echo=FALSE}
globe
```

Sure looks like aliens like the US...

Tools: R, LibreOffice Calc

Data Source: [UFO Sightings around the world](https://www.kaggle.com/camnugent/ufo-sightings-around-the-world)