---
title: "US Wildfires by County"
author: "Russell Szumski"
date: 2018-01-30
categories: ["R"]
tags: ["highcharter", "reshape2", "RSQLite"]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(htmltools)
df <- read.csv('Wildfires/Wildfires.csv')
df_melt <- df
df$fips <- sprintf("%05s", df$fips)
df$fips <- gsub(' ', '0', df$fips)
df_fip <- read.csv('Wildfires/State FIPS code.csv')
```

I watched Only the Brave, a 2017 American biographical action drama film that tells the true story of the Granite Mountain Hotshots, which lead me to the Kaggle [1.88 Million US Wildfires](https://www.kaggle.com/rtatman/188-million-us-wildfires) data set, then on to one of the Kernels titled [Wildfire Exploratory Analysis](https://www.kaggle.com/captcalculator/wildfire-exploratory-analysis) which had some interesting analysis of wildfires in the US and a cool choropleth map of total US wildfires by County. That lead me to the highmaps section of [highcharter](http://jkunst.com/highcharter/highmaps.html), and finally I found my way to a 'adding motion to choropleths' blog post on [jkunst](http://jkunst.com/r/adding-motion-to-choropleths/) that replicated a visualization made by [New York Times](https://www.nytimes.com/interactive/2016/01/07/us/drug-overdose-deaths-in-the-us.html?_r=0) showing How the Epidemic of Drug Overdose Deaths Rippled Across America.

So I replicated that visualization replication using the 1.88 Million US Wildfires data set.

The results

```{r,echo=F}
htmltools::tags$iframe(src = 'https://rstudio-pubs-static.s3.amazonaws.com/358987_2e5378be208e416a85f32b893a5bb6d8.html', width = '800', height = '600', style="border:none;")
```

Here's a [link](https://rstudio-pubs-static.s3.amazonaws.com/358987_2e5378be208e416a85f32b893a5bb6d8.html) to a full page view.

And here is how I did it.

The data is in a SQL database but because it will fit into RAM let’s get the data from the database and extract it into a data frame.

```{r, eval=FALSE}
library(RSQLite)
conn <- dbConnect(SQLite(), 'FPA_FOD_20170508.sqlite')
df <- tbl(conn, "Fires") %>% collect()
dbDisconnect(conn)
```

Looking at the data the `FIPS_CODE` has the FIPS code for the county but not the State so let's get that from the [State FIPS Code Listing](https://www.mcc.co.mercer.pa.us/dps/state_fips_code_listing.htm)

```{r, eval=FALSE}
library(rvest)
library(dplyr)

url <- 'https://www.mcc.co.mercer.pa.us/dps/state_fips_code_listing.htm'
webpage <- read_html(url)

STATE <- html_nodes(webpage, 'center a') %>%
  html_text()
FIPS_STATE <- html_nodes(webpage, 'tr+ tr td:nth-child(5) , tr+ tr td:nth-child(2)') %>%
  html_text() %>%
  .[-55]

df_state_fips <- as.data.frame(cbind(STATE, FIPS_STATE)) %>% mutate_if(is.factor, as.character)
```

```{r, echo=FALSE}
head(df_fip)
```


Select the needed variables, join the state FIPS codes this the data frame, unite the FIP codes for the states and counties into a new `fips` variable, calculate the number of fires for each group of `fips` and `FIRE_YEAR` (called `value`), and rename `FIRE_YEAR` to `year`.

```{r, eval=F}
df_ <- df %>%
  select(FIPS_CODE, FIRE_YEAR, STATE) %>%
  na.omit(.)

df_ <- left_join(df_state_fips, df_) %>%
  unite(fips, FIPS_STATE, FIPS_CODE, sep = '')

df_ <- df_ %>%
  group_by(fips, FIRE_YEAR) %>%
  summarize(value = n()) %>%
  rename(year = FIRE_YEAR)
```

Use the [reshape2](https://github.com/hadley/reshape) package to reshape and aggregate the data. 

First using `dcast` to take the data from long-format and casts it into wide-format by taking
`fips` and `value` as the ID variables (we want a column for each) thus giving a `variable` that describes the measured variables.

Then using `melt` we'll get values of `year` for each `fips` and `value`.

Final we'll summarize the sum of the `value` for the groups of `fips` and `vaiable`

```{r, eval=F}
df_cast <- dcast(df_, fips+value~year)
df_cast <- df_cast[,-27]

df_melt <- melt(df_cast, id = c('fips','value'))
df_melt <- df_melt[,-2]

## replace NA with 0
df_melt <- df_melt %>%
  mutate(value = ifelse(is.na(value), 0, value))

df_melt <- df_melt %>% 
  group_by(fips, variable) %>%
  summarize(value = sum(value))
```

```{r, echo=FALSE}
head(df_melt)
```

Prepare the data as the motion plugin requires the data in sequences in a list.

```{r, eval=F}
ds <- df_melt %>% 
  group_by(fips) %>% 
  do(item = list(
    fips = first(.$fips),
    sequence = .$value,
    value = first(.$value))) %>% 
  .$item
```

Building the plot.

```{r, eval=F}
data("uscountygeojson")

highchart(type = "map") %>%
  hc_add_series(data = ds,
                name = "Wildfires per year",
                mapData = uscountygeojson,
                joinBy = "fips",
                borderColor = '#000000',
                borderWidth = 0.25) %>% 
  hc_colorAxis(minColor = "#ffeda0", maxColor = "#f03b20") %>%  
  hc_title(text = "US Wildfires by County") %>% 
  hc_legend(layout = "vertical", reversed = TRUE,
            floating = TRUE, align = "right") %>%
  hc_motion(
    enabled = TRUE,
    axisLabel = "year",
    labels = sort(unique(df_melt$variable)),
    series = 0,
    updateIterval = 50,
    updateIterval = 50,
    magnet = list(
      round = "floor",
      step = 0.1
    )
  )
```
